DaLi: Database as a Library

### 2017-02-19 paper/intro.tex
Intro points added.

>>>>

Points:
\begin{itemize}
  \item Language is often considered the pivot of the thought process.
  Unsurprisingly, programmers are conditioned to reason in terms of
  the language-level abstractions. However, reality of underlying
  systems often force programmers to think in terms of low-level
  abstractions that are counter-intuitive from the language
  standpoint. Even functional programmers are often exposed to this
  rather unfortunate state of affairs.

  \item Consider, for example, an OCaml programmer who wishes to
  manage an ordered set of elements in a binary tree, while using a
  hashtable to quickly obtain references into the tree. While the
  programmer enjoys high-level abstractions and safety guarantees of
  OCaml insofar as the program only accesses in-memory data, she has
  to let go of this convenience if the application requires data
  persistence. She has to now re-engineer her application around the
  data model enforced by the underlying storage engine (e.g.,
  relational data model, or key-value data model), complicating
  programming, and potentially compormising the language safety
  guarantees in the process. For instance, while OCaml guarantees the
  safety of all references, references between tables in a database
  are unsafe, and have been known to lead to bugs.

  \item Add replication to persistence, and the problems grow
  manifold. The impossibility of obtaining strong consistency
  guarantees without compromising scalability/availability under
  decentralized replication means that applications can now exhibit
  anomalous behaviors that are often incomprehensible to the
  programmer. The standard artifacts of replication made available to
  the programmer, namely mutable key-value pairs, expose the underlying
  system reality in its full ugliness. While the programmer may manage
  to reason about the contents of a key-value cell with much effort,
  the reasoningly unfortunately does not compose; each key-value pair
  can be manipulated independently, which makes it almost impossible
  to obtain guarantees over a collection of data items in the same
  vein as a high-level data structure (e.g., a binary search tree).
  CRDTs are an attempt to provide high-level semantics of replication
  for certain useful data types, however CRDTs do not come with a
  an associated replicated storage system and a programming framework,
  and programmers are again faced with the task of bridging the gap
  between the high-level semantics of the CRDT to the abstractions and
  guarantees offered by the underlying storage system. Moreover,
  standard data structures need significant re-engineering to be
  considered CRDTs (e.g., functions need to return \emph{effects} all
  of which need to commute), and programmers receive little or no tool
  support in this process. Moreover, CRDTs famously do not compose (in
  the sense that the invariants over the composition cannot be
  preserved), requiring programmers to resort to extraneous
  mechanisms with system-specific properties, such as transactions.

  \item In this paper we describe a library called \name that adresses
  the aforementioned problems. \name lets programmers obtain add
  persistence and fully-decentralized replication to their
  applications without ever having to leave the comforts of high-level
  reasoning in OCaml. Thus, in keeping with the long tradition of
  Functional Programming, programmers are \emph{completely} shielded
  from the reality of the underlying system implementation. This
  elevation from the low-level reasoning to the high-level reasoning
  concretely manifests as reasoning in terms of \emph{mergeable} data
  types instead of \emph{replicated} data types; replication being a
  system-level concern.

\end{itemize}
